Estimating dark matter halo concentration using the integrated mass profile

### 2015-11-11 paper/paper.tex
puliendo texto del paper

\begin{abstract}
We present a new algorithm to estimate the concentration of N-body
dark matter halos.
The method uses the integrated mass profile to find the optimal
concentration value. 
Its main advantage is that it uses the full particle information
without any binning, making it realiable in  cases when low numerical
resolution becomes a limitation for other methods.  
Additionally, the algorithm is straightforward to implement.  
We test the performance of this method by estimating halo
concentration both on mock and N-body halos. 
We compare these results against two other methods:
maximum radial velocity measurements and radial particle binning to
fit the density profile. 
Tests on  the mock halos show that the accuracy of our method to recover input
concentrations varies with halo resolution, outperforming the other
two methods. For halos
sampled with $20$ particles our method recovers the input
concentration with $10\%$ accuracy, while for the maximum radial
velocity method it is on the order of $20\%$ 
and density profile methods fail at the $100\%$ level.
For halos with $2\times 10^4$ particles our method achieves a recovery
accuracy of $0.01\%$ while the velocity and density profile methods
achieve $0.1\%$ and $1\%$, respectively. 
We also measure the mass-concentration relationship on N-body
data. We find that within the probed mass range ($10^{12}\hMsun <
10^{14}\hMsun$) the three methods give consistent results within the
statistical uncertainties.   
We only find a small discrepancy at low masses, $M<10^{13}\hMsun$, where
our method yields lower median concentration values by $15\%-20\%$
compared to the velocity and density methods. 
From these results we believe that the new method deserves to be
included in a simulator's toolkit to probe the internal structure of
dark matter halos.
\end{abstract}

@@

%*************************************************************************
\section{Introduction}
\label{sec:introduction}
%*************************************************************************
In the current structure formation paradigm the properties of galaxies
are coupled to the evolution of their dark matter (DM) hosting halo.
This dependence has motivated the detailed study of dark matter halo
properties such as its formation history and internal structure. 
In particular, the sizes and dynamics of galaxies are determined by
the DM distribution inside a halo.

@@

The concentration of the NFW profile provides the conceptual
framework to study simulated dark matter halos as a function of
redshift and cosmological parameters.
In these studies 
\citep{Neto2007,Maccio2008,Duffy2008,Munoz2011,Prada2012,Ludlow2014} 
the results are summarized through the
mass-concentration relationship; that is the distribution of
concentration values at a fixed halo mass and redshift.
A big challenge in such numerical experiments is having a
reliable algorithm to estimate the concentration. 
Such an algorithm should provide unbiased results and must be robust
when applied to simulations of different numerical resolution.

>>>>

\begin{abstract}
We present a new algorithm to estimate the concentration of N-body
dark matter halos.
The method uses the integrated mass profile to find the optimal
concentration value. 
Its main advantage is that it uses the full particle information
without any binning, making it realiable in  cases when low numerical
resolution becomes a limitation for other methods.  
Additionally, the algorithm is straightforward to implement.  
We test the performance of this method by estimating halo
concentration both on mock and N-body halos. 
We compare these results against two other methods:
maximum radial velocity measurements and radial particle binning to
fit the density profile. 
Tests on  the mock halos show that the accuracy of our method to recover input
concentrations varies with halo resolution, outperforming the other
two methods. For halos
sampled with $20$ particles our method recovers the input
concentration with $10\%$ accuracy, while for the maximum radial
velocity method it is on the order of $20\%$ 
and density profile methods fail at the $100\%$ level.
For halos with $2\times 10^4$ particles our method achieves a recovery
accuracy of $0.01\%$ while the velocity and density profile methods
achieve $0.1\%$ and $1\%$, respectively. 
We also measure the mass-concentration relationship on N-body
data. We find that within the probed mass range ($10^{12}\hMsun <
10^{14}\hMsun$) the three methods give consistent results within the
statistical uncertainties.   
We only find a small discrepancy at low masses, $M<10^{13}\hMsun$, where
our method yields lower median concentration values by $15\%-20\%$
compared to the velocity and density methods. 
From these results we believe that the new method is promising tool
to probe the internal structure of dark matter halos.
\end{abstract}

@@

%*************************************************************************
\section{Introduction}
\label{sec:introduction}
%*************************************************************************
In the current structure formation paradigm the properties of galaxies
are coupled to the evolution of their dark matter (DM) hosting halo.
For instance, the sizes and dynamics of galaxies are determined by
the DM distribution inside a halo.
This dependence has motivated the detailed study of DM halo
properties such as its formation history and internal structure.

@@

The concentration of the NFW profile provides the conceptual
framework to study simulated dark matter halos as a function of
redshift and cosmological parameters.
In these studies 
\citep{Neto2007,Maccio2008,Duffy2008,Munoz2011,Prada2012,Ludlow2014} 
the results are summarized through the
mass-concentration relationship; that is the distribution of
concentration values at a fixed halo mass and redshift.
The success of such numerical experiments rests on a
reliable algorithm to estimate the concentration. 
Such an algorithm should provide unbiased results and must be robust
when applied to simulations of different numerical resolution.
